>

### 2025-07-10 docs/article.qmd
Render the revision of the methods and results

format:
  elsevier-pdf:
    keep-tex: true
    include-in-header:
      file: _header.tex
    journal:
      formatting: review
      cite-style: authoryear
    output-file: iadb-network-analysis.pdf
  docx:
    pandoc-options: ["--mathml"]
    output-file: iadb-network-analysis.docx

@@

Mathematically, a graph $\mathcal{G}$ is a pair $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, where $\mathcal{V}$ is the set of nodes and $\mathcal{E}$ is the set of edges. In a DPN, the node set $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n\}$ represents a finite set of $n$ medications. Additionally, $\mathcal{E}$ is a finite set of edges, where each edge is a pair of nodes that corresponds to a co-prescription. In an undirected graph, the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections, $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [@estrada2012structure].

@@

tar_read(res_tbl_poly) |>
  dplyr::mutate(
    "year" = as.character(year),
    "mean_sd" = sprintf("%.2f [%.2f]", mean, sd),
    "median_iqr" = sprintf("%.2f (%.2f)", median, IQR)
  ) |>
  dplyr::select(type:medication, sum, mean_sd) |>
  tidyr::pivot_wider(names_from = type, values_from = c(sum, mean_sd)) |>
  dplyr::select(year, medication, sum_same, sum_multi, mean_sd_same, mean_sd_multi, dplyr::contains("total")) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    year = "Year",
    medication = "Medication",
    sum_same = "Same-class",
    sum_multi = "Multi-class",
    mean_sd_same = "Same-class",
    mean_sd_multi = "Multi-class",
    sum_total = "N",
    mean_sd_total = "Mean [SD]"
  ) |>
  flextable::add_header_row(
    values    = c("", "Population Level", "Per Person (Mean [SD])", "Total Co-prescription"),
    colwidths = c(2, 2, 2, 2),
    top       = TRUE
  ) |>
  flextable::footnote(
    i = 1,
    j = c(3, 5, 7),
    ref_symbols = c("1", "2", "3"),
    part = "header",
    value = flextable::as_paragraph(
      c(
        "The total occurrence of co-prescription on the population level/perscription-day",
        "The expected occurrence of co-prescription per person/perscription-day",
        "The overall number of medications a patient uses, including those prescribed for multiple psychiatric symptom clusters as well as for comorbid non-psychiatric conditions",
        "Prescription day is defined as the number of dispensed medication per day",
        "Multi-class co-prescription is defined as the co-prescription of medications from different ATC classification"
      )
    )
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1:2, align = "left", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::width(j = c(1:8), c(1, rep(1.8, 7)), unit = "cm") |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

@@

res_recon_stat |>
  dplyr::arrange(-eigen_mean) |>
  dplyr::mutate(
    "mean_sd_eigen"     = sprintf("%.2e [%.2e]", eigen_mean, eigen_sd),
    "median_IQR_eigen"  = sprintf("%.2e [%.2e]", eigen_median, eigen_IQR),
    "mean_sd_dose"      = sprintf("%.2e [%.2e]", dose_mean, dose_sd),
    "median_IQR_dose"   = sprintf("%.2e [%.2e]", dose_median, dose_IQR),
    "mean_sd_weight"    = sprintf("%.2e [%.2e]", weight_mean, weight_sd),
    "median_IQR_weight" = sprintf("%.2e [%.2e]", weight_median, weight_IQR),
    "range_dose"        = sprintf("%.2e - %.2e", dose_min, dose_max),
    "range_weight"      = sprintf("%.2e - %.2e", weight_min, weight_max)
  ) |>
  subset(select = c(group, n_claim, mean_sd_eigen, mean_sd_dose, mean_sd_weight)) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    group            = "",
    n_claim          = "Dispensing",
    mean_sd_eigen    = "Centrality",
    mean_sd_dose     = "DDD",
    mean_sd_weight   = "Weight"
  ) |>
  flextable::add_header_row(
    values    = c("", "Mean [SD]"),
    colwidths = c(2, 3),
    top       = TRUE
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1, align = "left", part = "body") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  flextable::autofit(add_w = 0, add_h = 0) |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

>>>>

format:
  elsevier-pdf:
    keep-tex: true
    keep-md: true
    include-in-header:
      file: _header.tex
    journal:
      formatting: review
      cite-style: authoryear
    output-file: iadb-network-analysis.pdf
  docx:
    pandoc-options: ["--mathml"]
    output-file: iadb-network-analysis.docx

@@

Mathematically, a simple undirected graph $\mathcal{G}$ is a pair $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, where $\mathcal{V} = \left(1, \dots, n \right)$ is the set of nodes and $\mathcal{E} = \left(\{j, k\} \ | \ j, k \in \mathcal{V}, \ j \neq k \right)$ is the finite set of edges with $\{j, k\} = \{k, j\}$. The adjacency matrix $A \in \{0, 1\}^{n \times n}$ of a simple undirected graph $\mathcal{G}$ with $\mathcal{V}$ nodes is a square symmetric matrix of dimension $n \times n$ defined as @eq-graph-theory, with $A_{jk} = A_{kj}$ for all $j, k \in \mathcal{V}$ where $j \neq k$. For each day $t = \left(1, \dots, T \right)$ and each individual $p = \left(1, \dots, P \right)$, the adjacency matrix represented a simple undirected graph.

$$
A_{jk} = 
\begin{cases}
1 & \text{if } \{j, k\} \in \mathcal{E} \\
0 & \text{otherwise}
\end{cases}
$$ {#eq-graph-theory}

In a DPN, the node set $\mathcal{V} = \{\mathcal{v}_1, \mathcal{v}_2, \dots, \mathcal{v}_n\}$ represents a finite set of $n$ medications. Additionally, $\mathcal{E}$ is a finite set of edges, where each edge is a pair of nodes that corresponds to a co-prescription. In an undirected graph, the order of $i, j$ does not matter. Typically, undirected graphs do not contain self connections, $\{\mathcal{v}_i, \mathcal{v}_i\} \notin \mathcal{E}$ [j].

@@

tar_read(res_tbl_poly) |>
  dplyr::mutate(
    "year" = as.character(year),
    "mean_sd" = sprintf("%.2f [%.2f]", mean, sd),
    "median_iqr" = sprintf("%.2f (%.2f)", median, IQR)
  ) |>
  dplyr::select(type:medication, sum, mean_sd) |>
  tidyr::pivot_wider(names_from = type, values_from = c(sum, mean_sd)) |>
  dplyr::select(year, medication, sum_same, sum_multi, mean_sd_same, mean_sd_multi, dplyr::contains("total")) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    year = "Year",
    medication = "Medication",
    sum_same = "Same-class",
    sum_multi = "Multi-class",
    mean_sd_same = "Same-class",
    mean_sd_multi = "Multi-class",
    sum_total = "N",
    mean_sd_total = "Mean [SD]"
  ) |>
  flextable::add_header_row(
    values    = c("", "Population Level", "Per Person (Mean [SD])", "Total Co-prescription"),
    colwidths = c(2, 2, 2, 2),
    top       = TRUE
  ) |>
  flextable::footnote(
    i = 1,
    j = c(3, 5, 7),
    ref_symbols = c("1", "2", "3"),
    part = "header",
    value = flextable::as_paragraph(
      c(
        "The total occurrence of co-prescription on the population level/prescription-day.",
        "The expected occurrence of co-prescription per person/prescription-day.",
        "The overall number of medications a patient uses, including those prescribed for multiple psychiatric symptom clusters as well as for comorbid non-psychiatric conditions.",
        "Prescription day is defined as the number of dispensed medication per day.",
        "Multi-class co-prescription is defined as the co-prescription of medications from different ATC classification."
      )
    )
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1:2, align = "left", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::width(j = c(1:8), c(1, rep(1.8, 7)), unit = "cm") |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

@@

res_recon_stat |>
  dplyr::arrange(-eigen_mean) |>
  dplyr::mutate(
    "mean_sd_eigen"     = sprintf("%.2e [%.2e]", eigen_mean, eigen_sd),
    "median_IQR_eigen"  = sprintf("%.2e [%.2e]", eigen_median, eigen_IQR),
    "mean_sd_dose"      = sprintf("%.2e [%.2e]", dose_mean, dose_sd),
    "median_IQR_dose"   = sprintf("%.2e [%.2e]", dose_median, dose_IQR),
    "mean_sd_weight"    = sprintf("%.2e [%.2e]", weight_mean, weight_sd),
    "median_IQR_weight" = sprintf("%.2e [%.2e]", weight_median, weight_IQR),
    "range_dose"        = sprintf("%.2e - %.2e", dose_min, dose_max),
    "range_weight"      = sprintf("%.2e - %.2e", weight_min, weight_max)
  ) |>
  subset(select = c(group, n_claim, mean_sd_eigen, mean_sd_dose, mean_sd_weight)) |>
  flextable::flextable() |>
  flextable::set_header_labels(
    group            = "",
    n_claim          = "Dispensing",
    mean_sd_eigen    = "Centrality",
    mean_sd_dose     = "DDD",
    mean_sd_weight   = "Weight"
  ) |>
  flextable::add_header_row(
    values    = c("", "Mean [SD]"),
    colwidths = c(2, 3),
    top       = TRUE
  ) |>
  flextable::footnote(
    i = 2,
    j = c(4, 5),
    ref_symbols = c("1", "2"),
    part = "header",
    value = flextable::as_paragraph(
      c(
        "Defined Daily Dose, representing the assumed average maintenance dose per day for a drug used for its main indication in adults (WHO definition).",
        "Edge weight assigned to each medication class in the drug prescription network, reflecting the strength of co-prescription connections (see Methods: Data pre-processing to build the data matrix)."
      )
    )
  ) |>
  flextable::align(align = "center", part = "all") |>
  flextable::align(j = 1, align = "left", part = "all") |>
  #flextable::font(fontname = "Times", part = "all") |>
  flextable::fontsize(size = 7, part = "all") |>
  flextable::autofit(add_w = 0, add_h = 0) |>
  flextable::set_table_properties(opts_pdf = list("arraystretch" = 1, "tabcolsep" = 0))

### 2025-07-10 docs/exec-summary/_preprocessing.qmd
Render the revision of the methods and results

At the individual level, each patient's daily prescription record at time $t$ was encoded into a square matrix $A_t^{(i)}$, where rows and columns represent medication classes according to the ATC classification. An off-diagonal entry $A_{t_{jk}}^{(i)} = 1$ indicates that medication classes $i$ and $k$ were co-prescribed on the same day for individual $i$, while a diagonal entry $A_{t_{jj}}^{(i)} = 1$ captures the presence of a single prescription of medication $j$. To derive the population-level DPN, all individual-level daily prescription matrix $A^{(i)}$ were aggregated by computing their element-wise sum.

@@

The population matrix $A_t$ captures the frequency of co-prescriptions (off-diagonal) and single prescriptions (diagonal) across the dataset at the specified time $t$. Each element $A_{t_{jk}}$ in the matrix was interpreted as the number of patients being prescribed with medication classes $j$ and $k$ on the same day $t$. $A_t$ was represented as a weighted undirected graph, where each node corresponds to a medication class and edges reflect co-prescription frequencies. In the original work of DPN, all defined daily dose (DDD) were assumed equal to 1, allowing for applying raw counts as the edge weights. However, since not all medications have DDD = 1, this assumption could lead to inaccurate representations. To enhance the accuracy of edge weights, we accounted for the heterogeneity in DDD. The weights were adjusted using the density of a normal distribution. The expected value ($\mu$) was set to the base weight $\omega_B$ of 1 to prioritize DDD = 1, and the standard deviation ($\sigma$) was set to $\frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1.

$$\omega = \frac{1}{\frac{\omega_B}{3} \sqrt{2 \pi}} e^{-\frac{1}{2}\left(\frac{DDD - \omega_B}{\omega_B/3}\right)^2}$$ {#eq-weighted-DDD}

@@

>>>>

At the individual level, each patient's daily prescription record at time $t$ was encoded into a square matrix $A_t^{(i)}$, where rows and columns represent medication classes according to the ATC classification. An off-diagonal entry $A_{t_{jk}}^{(i)} = 1$ indicates that medication classes $j$ and $k$ were co-prescribed on the same day for individual $i$, while a diagonal entry $A_{t_{jj}}^{(i)} = 1$ captures the presence of a single prescription of medication $j$. To derive the population-level DPN, all individual-level daily prescription matrix $A^{(i)}$ were aggregated by computing their element-wise sum.

@@

The population matrix $A_t$ captures the frequency of co-prescriptions (off-diagonal) and single prescriptions (diagonal) across the dataset at the specified time $t$. Each element $A_{t_{jk}}$ in the matrix was interpreted as the number of patients being prescribed with medication classes $j$ and $k$ on the same day $t$. $A_t$ was represented as a weighted undirected graph, where each node corresponds to a medication class and edges reflect co-prescription frequencies. In the original work of DPN, all defined daily dose (DDD) were assumed equal to 1, allowing for applying raw counts as the edge weights. However, since not all medications have DDD = 1, this assumption could lead to inaccurate representations. To enhance the accuracy of edge weights, we introduced a weighting function to adjust the contribution of each co-prescription. Specifically, each co-prescribed medication class was assigned a weight $\omega_j$ based on its DDD value using a Gaussian kernel centered at the baseline weight $\omega_B = 1$, i.e., the expected DDD, with standard deviation $\sigma = \frac{1}{3}$ to allow for a gradual reduction in weight as DDD deviates from 1.

$$\omega_j = \frac{1}{\frac{\omega_B}{3} \sqrt{2 \pi}} e^{-\frac{1}{2}\left(\frac{DDD - \omega_B}{\omega_B/3}\right)^2}$$ {#eq-weighted-DDD}

The edge weight between medication classes $j$ and $k$ was computed as the average of their individual weights:

@@

This weighting approach scaled the edge weight that deviate substantially from the standard $\omega_B = 1$, improving comparability across medication classes. To remain consistent with standard network representations and avoid confounding self-loops, the diagonal elements were omitted.
